---
title: "Reading Data From The Web"
output: github_document
date: "2022-10-13"
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)
library(httr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Extracting Tables

### NSDUH Dataset

Using `read_html` in the rvest package, we can read in the html from a webpage.

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

drug_use_html
```

```{r}
drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

`html_table` pulls out all tables from the webpage. It's not giving a dataframe, but a list (a special structure for storing information), and so we can use `first` to pull out the first item from the list and output it as a table.

The “note” at the bottom of the table appears in every column in the first row. We need to remove that using `slice(-1)`.

## CSS Selectors 

### Star Wars movies

Suppose we’d like to scrape the data about the Star Wars Movies from the IMDB page. The first step is the same as before – we need to get the HTML.

```{r}
swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```


The information isn’t stored in a handy table, so we’re going to isolate the CSS selector for elements we care about. A bit of clicking around gets me something like below.



