---
title: "Strings and Factors"
output: github_document
date: "2022-10-18"
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## String vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff") #Detect "jeff" in your vector of strings
str_detect(string_vec, "a") # Looks for anything in string that contains "a"

str_replace(string_vec, "jeff", "Jeff")


string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think") # Find something that contains "i think"
str_detect(string_vec, "^i think") # Find something that STARTS with "i think"
str_detect(string_vec, "i think$") # Find something that ENDS with "i think"
```

You can designate a list of characters that will count as a match.

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"Bush") # Only detects Bush with capital B
str_detect(string_vec,"[Bb]ush") # Detects bush with a capital or lowercase b
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][A-Z]") 
str_detect(string_vec, "^[0-9][a-zA-Z]") 
```

The character `.` matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # . means anything can be between a "7" and "11", but must have something

```

Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. If you want to search for these, you have to indicate they’re special using `\`. Unfortunately, `\` is also special, so things get weird.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "[") # Error message expects you to put something in the bracket and indicate a range

str_detect(string_vec, "\\[") # Put \\ to search for a special character

str_detect(string_vec, "\\[0-9]") 

```

## Why factors are weird

Factors are the way to store categorical variables in R. They can take on specific levels (e.g. male and female) which are usually presented as characters but are, in fact, stored by R as integers. These integer values are used by functions throughout R – in making plots, in organizing tables, in determining the “reference” category – but most of the time are hidden by easier-to-read character string labels. This close relationship to strings, when in fact there is a lot of added structure, is why factors can be so confusing.

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

as.numeric(factor_vec) # Male = 2 

factor_vec = fct_relevel(factor_vec, "male") # Relevel male as the first level

as.numeric(factor_vec)
```

The previous code also illustrates coersion: forcing a variable from one type (e.g. factor) to another (e.g. numeric). Understanding how R coerces variables is important, because it sometimes happens unintentionally and can break your code or impact your analyses.

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)
```

There are a few steps we need to implement to tidy these data.

```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

We used stringr and regular expressions a couple of times above:

* in `separate`, we split age and year at the open parentheses using `"\\("`
* we stripped out the close parenthesis in `mutate`
* to remove character superscripts, we replaced any character using `"[a-c]$"`

Let’s quickly visualize these data for the 12-17 age group; to make the plot readable, we’ll treat `State` as a factor are reorder according to the median `percent` value.
